---
title: "Late Delivery Model"
subtitle: "Skill 3 — Logistic regression risk screening"
---

```{r model-setup, echo=FALSE}
source("R/website_helpers.R")
silent_run({
  source("R/model_low_reviews.R")
  res <- train_predictive_model("data/processed/order_level_analysis.rds")
})
late_prev <- as.numeric(res$class_balance_pct["1"])
if (is.na(late_prev)) {
  late_prev <- 100 * mean(readRDS("data/processed/order_level_analysis.rds")$late_delivery, na.rm = TRUE)
}
cm <- res$confusion_matrix
coef_join <- merge(
  res$odds_ratios,
  res$coefficients[, c("term", "p_value")],
  by = "term",
  all.x = TRUE
)
top_or <- coef_join[!grepl("^\\(Intercept\\)|^purchase_month|^primary_payment_type", coef_join$term), ]
top_or <- top_or[order(abs(log(top_or$odds_ratio)), decreasing = TRUE), ]
top_or_main <- head(top_or, 6)
```

<div class="question-card">
<div class="card-title">Business Question</div>
Which available order characteristics are associated with whether an order is delivered late?
</div>

```{r model-kpis, echo=FALSE, results='asis'}
cat(kpi_row(c(
  kpi_card("graph-up", "ROC AUC", sprintf("%.3f", res$roc_auc), "Test-set discrimination", "blue"),
  kpi_card("bullseye", "Sensitivity", sprintf("%.1f%%", 100 * res$sensitivity), "Recall on test set", "green"),
  kpi_card("filter", "Precision", sprintf("%.1f%%", 100 * res$precision), "Positive predictive value", "orange"),
  kpi_card("percent", "Late Prevalence", sprintf("%.2f%%", late_prev), "Analysis sample", "red")
)))
```

<div class="two-col-grid">

<div class="dash-panel chart-panel">
<div class="panel-header"><span class="panel-icon"><i class="bi bi-bezier2"></i></span> ROC Curve</div>
<div class="panel-body">

```{r model-roc-plot, echo=FALSE, fig.align="center", fig.height=3.8}
roc <- read.csv("output/model/roc_curve.csv")
roc <- roc[order(roc$threshold, decreasing = TRUE), ]
idx <- unique(c(1, seq(1, nrow(roc), length.out = min(250, nrow(roc))), nrow(roc)))
roc_plot <- roc[idx, ]
plot(
  roc_plot$fpr, roc_plot$tpr,
  type = "l", col = "#2563EB", lwd = 2,
  xlab = "False Positive Rate", ylab = "True Positive Rate",
  main = sprintf("ROC Curve (AUC = %.3f)", res$roc_auc)
)
abline(0, 1, lty = 2, col = "#9CA3AF")
```

</div>
</div>

<div class="dash-panel">
<div class="panel-header"><span class="panel-icon"><i class="bi bi-grid-3x3"></i></span> Confusion Matrix (Test Set)</div>
<div class="panel-body">

```{r model-confusion-grid, echo=FALSE, results='asis'}
cat(sprintf(
  paste0(
    '<div class="confusion-grid">',
    '<div></div><div class="cg-header">Pred On-Time</div><div class="cg-header">Pred Late</div>',
    '<div class="cg-label">Actual On-Time</div><div class="cg-cell cg-tn">%s</div><div class="cg-cell cg-fp">%s</div>',
    '<div class="cg-label">Actual Late</div><div class="cg-cell cg-fn">%s</div><div class="cg-cell cg-tp">%s</div>',
    '</div>'
  ),
  format(cm$count[1], big.mark = ","),
  format(cm$count[2], big.mark = ","),
  format(cm$count[3], big.mark = ","),
  format(cm$count[4], big.mark = ",")
))
```

<p style="text-align:center;font-size:0.8rem;color:#6B7280;margin-top:0.75rem;">
Threshold: **`r sprintf("%.3f", res$classification_threshold)`** | Accuracy **`r sprintf("%.1f%%", 100 * res$accuracy)`** | F1 **`r sprintf("%.1f%%", 100 * res$f1_score)`**
</p>

</div>
</div>

</div>

<div class="dash-panel chart-panel">
<div class="panel-header"><span class="panel-icon"><i class="bi bi-bar-chart-line"></i></span> Top Predictor Odds Ratios</div>
<div class="panel-body">

```{r model-or-chart, echo=FALSE, fig.align="center", fig.height=4.2}
or_vals <- top_or_main$odds_ratio
labels <- predictor_label(top_or_main$term)
ord <- order(or_vals)
or_vals <- or_vals[ord]
labels <- labels[ord]
xmax <- max(or_vals) * 1.08
par(mar = c(4, 11, 3, 2))
barplot(
  or_vals,
  horiz = TRUE,
  names.arg = labels,
  las = 1,
  col = ifelse(or_vals > 1, "#DC2626", "#2563EB"),
  border = NA,
  main = "Top Predictors by Odds Ratio",
  xlab = "Odds Ratio",
  xlim = c(0, xmax),
  cex.names = 0.82
)
abline(v = 1, lty = 2, col = "#9CA3AF", lwd = 1.5)
```

</div>
</div>

<div class="dash-panel">
<div class="panel-header"><span class="panel-icon"><i class="bi bi-sliders"></i></span> Model Summary</div>
<div class="panel-body">

| Item | Value |
| :--- | :--- |
| Analysis sample | `r format(res$analysis_sample_size, big.mark = ",")` orders |
| Training / test split | `r format(res$train_size, big.mark = ",")` / `r format(res$test_size, big.mark = ",")` |
| Threshold selection | Training-only Youden's J |
| Selected threshold | `r sprintf("%.3f", res$classification_threshold)` |
| Specificity | `r sprintf("%.1f%%", 100 * res$specificity)` |

**Key predictors used in the model:**

- Order value
- Freight cost
- Number of items
- Number of sellers
- Estimated delivery window
- Maximum payment installments
- Purchase month
- Primary payment type

<p class="muted-note">Skewed monetary variables were log-transformed before modeling.</p>

</div>
</div>

<div class="dash-panel">
<div class="panel-header"><span class="panel-icon"><i class="bi bi-list-check"></i></span> Key Predictors</div>
<div class="panel-body">

```{r model-or-table, echo=FALSE}
or_display <- transform(top_or_main, odds_ratio = round(odds_ratio, 4), p_value = format_pvalue(p_value))
knitr::kable(
  or_display[, c("term", "odds_ratio", "p_value")],
  col.names = c("Predictor", "Odds Ratio", "p-value"),
  row.names = FALSE
)
```

</div>
</div>

<div class="insight-card">
<div class="card-title">Practical Interpretation</div>

- Accuracy near **`r sprintf("%.0f%%", 100 * res$accuracy)`** is misleading under class imbalance.
- Precision is only **`r sprintf("%.1f%%", 100 * res$precision)`**; use as a **screening tool**, not an automatic decision rule.
- Association does not prove causation. Log-transformed monetary odds ratios are not one-real changes.

</div>

<div class="warning-card">
<div class="card-title">Limitation</div>
Late deliveries are imbalanced, precision remains low, and undelivered or canceled orders are excluded from the modeling sample.
</div>

<details class="dash-details">
<summary>Full Model Coefficients</summary>
<div class="details-body">

```{r model-full-or-table, echo=FALSE}
full_or <- coef_join
full_or$odds_ratio <- round(full_or$odds_ratio, 4)
full_or$or_ci_lower <- round(full_or$or_ci_lower, 4)
full_or$or_ci_upper <- round(full_or$or_ci_upper, 4)
full_or$p_value <- format_pvalue(full_or$p_value)
knitr::kable(
  full_or[, c("term", "odds_ratio", "or_ci_lower", "or_ci_upper", "p_value")],
  col.names = c("Predictor", "Odds Ratio", "95% CI Lower", "95% CI Upper", "p-value"),
  row.names = FALSE
)
```

</div>
</details>

<details class="dash-details">
<summary>Method, population, and class distribution</summary>
<div class="details-body">

Delivered orders with valid estimated and actual delivery dates, complete predictors, and non-`not_paid` payment type.

| Partition | On-Time (0) | Late (1) |
| :--- | ---: | ---: |
| Training | `r format(res$train_distribution[1], big.mark = ",")` | `r format(res$train_distribution[2], big.mark = ",")` |
| Test | `r format(res$test_distribution[1], big.mark = ",")` | `r format(res$test_distribution[2], big.mark = ",")` |

`r res$threshold_reason`

</div>
</details>
